---
title: "Descriptive Analysis UKHLS"
author: "Emir Zecovic"

format:
  html:
    theme: spacelab
    embed-resources: true
    smooth-scroll: true
    toc: true
    toc-location: left
    toc-depth: 6
    toc-title: Contents

number-sections: false

comments:
  hypothesis: false

editor: visual
---

```{r, include=FALSE}
# Setup ----
## Packages ----
# if ("convenience" %in% rownames(installed.packages()) ==F) {
#   devtools::install_github("ratsupaltuf/convenience", force=T)
# }

alabs<- function(x, kable=TRUE, extended=T){
  tb<- tibble(labels=names(attributes(x)$labels), values=attributes(x)$labels)
  
  if(extended==T) {
    lab<- attributes(x)$label
    cl<- class(x)
    u<- unique(x)
    
    if(kable==T) {
      
      cat(lab, "\n",
          "Class:",cl, "\n",
          "Unique values:", u, "\n",
          "\n"
      )
      print(knitr::kable(tb, caption="Labels"))
      
    }  else {
      cat(lab, "\n",
          "Class:",cl, "\n",
          "Unique values:", u, "\n",
          "\n"
      )
      print(tb)
    }
  } else if (kable==T) {
    
    kable(tb)
  }  else {
    tb
  }
}

packages <- c("tidyverse", "haven", "pastecs", "datawizard", #"convenience",
              "ggplot2", "ggalluvial", "ggthemes", "viridis", "ggrepel",
              "sjPlot", "lme4", "knitr", "kableExtra", "gt", "survey",
              "stringr", "flextable", "officer", "sf", "plm", "stargazer",
              "patchwork", "tidytext", "sjlabelled")
install.packages(setdiff(packages, rownames(installed.packages())))
suppressMessages(lapply(packages, library, character.only = TRUE, quietly=T))
rm(packages)
options(max.print=10000)

## Load ----
uk <- readRDS("C:/Users/Emir  PC/Desktop/PhD/Paper1/PAIRS/data/uk.rds")
#uk <- readRDS("~/PAIRS/data/uk.rds")
```

## Means-tested benefit coding

```{r, eval=FALSE, echo=TRUE}
uk <- uk %>%
  mutate(
    benefit_MT = case_when(
      
      # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # 
      # WAVES 1-5: RECEIPT
      # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # #
      
      wave %in% 1:5 &
        (
          # JSA/unemp. benefits
          bt_benefit_unemp == 1 |
            unemp_benefit_JSA == 1 |
            inct_benefit_JSA == 1 |
            
            # Income Support
            bt_benefit_IS == 1 |
            
            # Housing Benefit
            hou_benefit_HB == 1 |
            inct_benefit_hous == 1 |
            
            # Working Tax Credit
            bt_benefit_WTC == 1 |
            tax_benefit_WTC == 1 |
            inct_benefit_WTC == 1 |
            
            # Council Tax Benefit 
            tax_benefit_CTS == 1 |
            
            # Universal Credit
            (wave %in% 3:5 &
               (
                 unemp_benefit_UC == 1 |
                   bt_benefit_UC == 1 |
                   hou_benefit_UC == 1 |
                   tax_benefit_UC == 1 |
                   dis_benefit_UC == 1 |
                   inct_benefit_UC == 1
               ))
        ) ~ 1,
      
      
      # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # 
      # WAVES 6-15: RECEIPT
      # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # #
      
      wave %in% 6:15 &
        (
          # JSA
          inc_benefit_JSA == 1 |
            inct_benefit_JSA == 1 |
            
            # Income Support
            inc_benefit_IS == 1 |
            
            # Housing Benefit
            oth_benefit_hous == 1 |
            inct_benefit_hous == 1 |
            
            # Working Tax Credit
            oth_benefit_WTC == 1 |
            inct_benefit_WTC == 1 |
            oth_benefit_IWC_lp == 1 |
            
            # Universal Credit
            inc_benefit_UC == 1 |
            inct_benefit_UC == 1
        ) ~ 1,
      
      
      # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # #
      # WAVES 1-5: NON-RECEIPT
      # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # #
      
      wave %in% 1:5 &
        (
          bt_benefit_unemp == 0 &
            bt_benefit_IS == 0 &
            bt_benefit_WTC == 0
        ) ~ 0,
      
      
      # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # #
      # WAVES 6-15: NON-RECEIPT
      # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # #
      
      wave %in% 6:15 &
        (
          inc_benefit_JSA == 0 &
            inc_benefit_IS == 0 &
            oth_benefit_hous == 0 &
            oth_benefit_WTC == 0 &
            inc_benefit_UC == 0
        ) ~ 0,
      
      
      # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # #
      # OTHERWISE: NA
      # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # #
      
      TRUE ~ NA_real_
    )
  )

```

## unweighted descriptives

```{r, echo=FALSE, error=FALSE}
uk %>%
  count(benefit_MT) %>%
  mutate(Percent = 100 * n / sum(n)) %>%
  round(2)
```



```{r, echo=FALSE, error=FALSE}
benefit_plot_data <- uk %>%
  select(wave, benefit_MT) %>%
  mutate(
    benefit_status = case_when(
      benefit_MT == 0 ~ "No",
      benefit_MT == 1 ~ "Yes",
      is.na(benefit_MT) ~ "Missing"
    ),
    benefit_status = factor(
      benefit_status,
      levels = c("No", "Yes", "Missing")
    )
  ) %>%
  count(wave, benefit_status)



ggplot(
  benefit_plot_data,
  aes(
    x = wave,
    y = n,
    fill = benefit_status
  )
) +
  geom_col() +
  scale_x_continuous(breaks = 1:15) +
  theme_tufte(base_size = 18) +
  labs(
    x = "Wave",
    y = "Frequency",
    fill = "Means-tested benefit receipt"
  ) +
  scale_fill_viridis_d(direction = -1)

```

## Weighted descriptives

### Benefit types


```{r, include=FALSE}
weights_cs <- read_dta("C:/Users/Emir  PC/Desktop/PhD/Paper1/PAIRS/data/weights_cs.dta")
#weights_cs <- read_dta("~/PAIRS/data/weights_cs.dta")

## Merge ----
uk <- uk %>%
  left_join(
    weights_cs,
    by = c("pidp", "wave")
  )

## Apply ----
### Missings ----
uk_cs <- uk %>%
  filter(
    !is.na(benefit_MT),
    weight_cs > 0,
    !is.na(psu),
    !is.na(strata)
  )


### Survey ----
options(survey.lonely.psu = "adjust")
uk_cs_w <- svydesign(
  ids = ~psu,
  strata = ~strata,
  weights = ~weight_cs,
  data = uk_cs,
  nest = TRUE
)

benefit_time_series <- uk_cs %>%
  group_by(wave) %>%
  summarise(
    n_unweighted = sum(benefit_MT == 1),
    n_weighted = sum(weight_cs * (benefit_MT == 1)),
    .groups = "drop"
  )

benefit_time_series



benefit_time_series <- uk_cs %>%
  group_by(wave) %>%
  summarise(
    n_unweighted = sum(benefit_MT == 1),
    n_weighted = sum(weight_cs * (benefit_MT == 1)),
    weighted_percent = 100 * weighted.mean(
      benefit_MT,
      w = weight_cs
    ),
    .groups = "drop"
  )
benefit_time_series
```


```{r, include=FALSE}
# Check ----
uk <- uk %>%
  mutate(
    # JSA
    check_JSA = case_when(
      wave %in% 1:5 &
        (
          bt_benefit_unemp == 1 |
            inct_benefit_JSA == 1 |
            unemp_benefit_JSA == 1
        ) ~ 1,
      
      wave %in% 6:15 &
        (
          inc_benefit_JSA == 1 |
            inct_benefit_JSA == 1
        ) ~ 1,
      
      TRUE ~ 0
    ),
    
    # Income Support
    check_IS = case_when(
      wave %in% 1:5 &
        bt_benefit_IS == 1 ~ 1,
      
      wave %in% 6:15 &
        inc_benefit_IS == 1 ~ 1,
      
      TRUE ~ 0
    ),
    
    # Housing Benefit
    check_HB = case_when(
      wave %in% 1:5 &
        (
          hou_benefit_HB == 1 |
            inct_benefit_hous == 1
        ) ~ 1,
      
      wave %in% 6:15 &
        (
          oth_benefit_hous == 1 |
            inct_benefit_hous == 1
        ) ~ 1,
      
      TRUE ~ 0
    ),
    
    # Working Tax Credit
    check_WTC = case_when(
      wave %in% 1:5 &
        (
          bt_benefit_WTC == 1 |
            tax_benefit_WTC == 1 |
            inct_benefit_WTC == 1
        ) ~ 1,
      
      wave %in% 6:15 &
        (
          oth_benefit_WTC == 1 |
            inct_benefit_WTC == 1
        ) ~ 1,
      
      TRUE ~ 0
    ),
    
    # Universal Credit
    check_UC = case_when(
      inc_benefit_UC == 1 |
        inct_benefit_UC == 1 |
        unemp_benefit_UC == 1 |
        bt_benefit_UC == 1 |
        hou_benefit_UC == 1 |
        tax_benefit_UC == 1 |
        dis_benefit_UC == 1 ~ 1,
      
      TRUE ~ 0
    )
  )
```


```{r, echo=FALSE}
benefit_components <- uk %>%
  filter(!is.na(benefit_MT)) %>%
  group_by(wave) %>%
  summarise(
    N_observed = n(),
    
    JSA = round(
      100 * mean(check_JSA == 1),
      1
    ),
    
    Income_Support = round(
      100 * mean(check_IS == 1),
      1
    ),
    
    Housing_Benefit = round(
      100 * mean(check_HB == 1),
      1
    ),
    
    WTC = round(
      100 * mean(check_WTC == 1),
      1
    ),
    
    UC = round(
      100 * mean(check_UC == 1),
      1
    ),
    
    Any_MT = round(
      100 * mean(benefit_MT == 1),
      1
    ),
    
    .groups = "drop"
  )

benefit_components %>%
  print(n = 15)

```

```{r, include=FALSE}
benefit_time_series_components <- uk %>%
  filter(
    weight_cs > 0,
    !is.na(psu),
    !is.na(strata)
  ) %>%
  group_by(wave) %>%
  summarise(
    JSA_unweighted = sum(check_JSA == 1, na.rm = TRUE),
    JSA_weighted = sum(weight_cs * (check_JSA == 1), na.rm = TRUE),

    IS_unweighted = sum(check_IS == 1, na.rm = TRUE),
    IS_weighted = sum(weight_cs * (check_IS == 1), na.rm = TRUE),

    HB_unweighted = sum(check_HB == 1, na.rm = TRUE),
    HB_weighted = sum(weight_cs * (check_HB == 1), na.rm = TRUE),

    WTC_unweighted = sum(check_WTC == 1, na.rm = TRUE),
    WTC_weighted = sum(weight_cs * (check_WTC == 1), na.rm = TRUE),

    UC_unweighted = sum(check_UC == 1, na.rm = TRUE),
    UC_weighted = sum(weight_cs * (check_UC == 1), na.rm = TRUE),

    .groups = "drop"
  )

benefit_time_series_components

benefit_plot_data <- benefit_time_series_components %>%
  select(
    wave,
    JSA_weighted,
    IS_weighted,
    HB_weighted,
    WTC_weighted,
    UC_weighted
  ) %>%
  pivot_longer(
    cols = -wave,
    names_to = "benefit",
    values_to = "n_weighted"
  ) %>%
  mutate(
    benefit = recode(
      benefit,
      JSA_weighted = "Jobseeker's Allowance",
      IS_weighted = "Income Support",
      HB_weighted = "Housing Benefit",
      WTC_weighted = "Working Tax Credit",
      UC_weighted = "Universal Credit"
    )
  )
```



```{r, echo=FALSE}
benefit_plot_data %>%
  ggplot(
    aes(
      x = wave,
      y = n_weighted,
      color = benefit,
      group = benefit
    )
  ) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 1:15) +
  theme_bw() +
  labs(
    x = "Wave",
    y = "Weighted number of benefit recipients",
    color = "Benefit",
    title = "Weighted Benefit Receipt by Wave"
  )
```






```{r, echo=FALSE}
benefit_time_series %>%
  ggplot(aes(x = wave, y = n_weighted)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 1:15) +
  theme_bw() +
  labs(
    x = "Wave",
    y = "Weighted number of means-tested benefit receipt",
    title = "Weighted Means-tested Benefit Receipt by Wave"
  )
```




```{r, echo=FALSE}
tab4_w <- uk %>%
  filter(
    !is.na(benefit_MT),
    weight_cs > 0
  ) %>%
  group_by(wave, benefit_MT) %>%
  summarise(
    n_unweighted = n(),
    n_weighted = sum(weight_cs),
    .groups = "drop"
  ) %>%
  group_by(wave) %>%
  mutate(
    prop = n_weighted / sum(n_weighted)
  ) %>%
  ungroup()

tab4_w %>%
  print(n = 100)
```


```{r, echo=FALSE}
tab4_w %>%
  ggplot(
    aes(
      x = wave,
      y = 100 * prop,
      fill = factor(
        benefit_MT,
        levels = c(0, 1),
        labels = c("No", "Yes")
      )
    )
  ) +
  geom_col(
    position = "dodge"
  ) +
  scale_x_continuous(
    breaks = 1:15
  ) +
  theme_bw() +
  labs(
    x = "Wave",
    y = "Weighted percent %",
    fill = "Means-tested benefit receipt"
  )
```





### Means-tested benefit receipt by wave

```{r, echo=FALSE, include=FALSE}
benefit_wave_w <- uk %>%
  filter(
    !is.na(benefit_MT),
    weight_cs > 0
  ) %>%
  group_by(wave, benefit_MT) %>%
  summarise(
    n_unweighted = n(),
    n_weighted = sum(weight_cs),
    .groups = "drop"
  ) %>%
  group_by(wave) %>%
  mutate(
    percent = 100 * n_weighted / sum(n_weighted)
  ) %>%
  ungroup() %>%
  mutate(
    benefit_MT = factor(
      benefit_MT,
      levels = c(0, 1),
      labels = c("No", "Yes")
    )
  )

benefit_wave_w %>%
  gt() %>%
  fmt_number(
    columns = c(n_weighted, percent),
    decimals = 1
  ) %>%
  cols_label(
    wave = "Wave",
    benefit_MT = "Means-tested benefit",
    n_unweighted = "Unweighted N",
    n_weighted = "Weighted N",
    percent = "Weighted %"
  ) %>%
  tab_header(
    title = "Weighted Means-tested Benefit Receipt by Wave"
  )
```


```{r, echo=FALSE}
benefit_wave_w %>%
  ggplot(
    aes(
      x = wave,
      y = percent,
      color = benefit_MT,
      group = benefit_MT
    )
  ) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(
    breaks = 1:15
  ) +
  theme_bw() +
  labs(
    x = "Wave",
    y = "Weighted percent %",
    color = "Means-tested benefit",
    title = "Weighted Means-tested Benefit Receipt by Wave"
  )
```

------------------------------------------------------------------------

### Means-tested benefit receipt by wave and labor force status

```{r, echo=FALSE}
benefit_lfstat_w <- uk %>%
  filter(
    !is.na(benefit_MT),
    !is.na(lfstat),
    weight_cs > 0
  ) %>%
  group_by(wave, lfstat, benefit_MT) %>%
  summarise(
    n_weighted = sum(weight_cs),
    .groups = "drop"
  ) %>%
  group_by(wave, lfstat) %>%
  mutate(
    percent = 100 * n_weighted / sum(n_weighted)
  ) %>%
  ungroup()

benefit_lfstat_w %>%
  filter(benefit_MT == 1) %>%
  ggplot(
    aes(
      x = wave,
      y = percent
    )
  ) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_wrap(~ lfstat) +
  scale_x_continuous(breaks = 1:15) +
  theme_bw() +
  labs(
    x = "Wave",
    y = "Weighted means-tested benefit receipt (%)",
    title = "Weighted Means-tested Benefit Receipt by Employment Status"
  )
```



```{r, echo=FALSE}
benefit_employment_w <- uk %>%
  mutate(
    employment = case_when(
      lfstat %in% c(
        "Full-time",
        "Part-time",
        "Self-employed"
      ) ~ "Employed (Full, part, self)",
      
      !is.na(lfstat) ~ "Not employed (Unemployed, Retired, Inactive)",
      
      TRUE ~ NA_character_
    )
  ) %>%
  filter(
    !is.na(benefit_MT),
    !is.na(employment),
    weight_cs > 0
  ) %>%
  group_by(wave, employment, benefit_MT) %>%
  summarise(
    n_weighted = sum(weight_cs),
    .groups = "drop"
  ) %>%
  group_by(wave, employment) %>%
  mutate(
    percent = 100 * n_weighted / sum(n_weighted)
  ) %>%
  ungroup()

benefit_employment_w %>%
  filter(benefit_MT == 1) %>%
  ggplot(
    aes(
      x = wave,
      y = percent
    )
  ) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_wrap(~ employment) +
  scale_x_continuous(breaks = 1:15) +
  theme_bw() +
  labs(
    x = "Wave",
    y = "Weighted means-tested benefit receipt (%)",
    title = "Weighted Means-tested Benefit Receipt by Employment Status"
  )
```





------------------------------------------------------------------------

### Means-tested benefit receipt by wave and paid work

```{r, echo=FALSE}
# uk %>%
#   count(wave, benefit_MT, job) %>%
#   na.omit() %>%
#   group_by(wave, job) %>%
#   mutate(percent = 100 * n / sum(n)) %>%
#   select(-n) %>%
#   print(n = 60)
```

```{r, echo=FALSE}
benefit_job_w <- uk %>%
  filter(
    !is.na(benefit_MT),
    !is.na(job),
    weight_cs > 0
  ) %>%
  group_by(wave, job, benefit_MT) %>%
  summarise(
    n_weighted = sum(weight_cs),
    .groups = "drop"
  ) %>%
  group_by(wave, job) %>%
  mutate(
    percent = 100 * n_weighted / sum(n_weighted)
  ) %>%
  ungroup()

benefit_job_w %>%
  filter(benefit_MT == 1) %>%
  ggplot(
    aes(
      x = wave,
      y = percent
    )
  ) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_wrap(~ job) +
  scale_x_continuous(breaks = 1:15) +
  theme_bw() +
  labs(
    x = "Wave",
    y = "Weighted means-tested benefit receipt (%)",
    title = "Weighted Means-tested Benefit Receipt by Paid Work Status"
  )
```




------------------------------------------------------------------------



### Trends by groups (unweighted)

**Wave comparison of mean relationship satisfaction by means-tested benefit receipt.**

```{r, echo=FALSE, error=FALSE}
# uk %>%
#   group_by(wave, benefit_MT) %>%
#   summarise(mean_rel_happy = mean(rel_happy, na.rm = T) %>%
#               round(2)) %>%
#   na.omit() %>%
#   spread(benefit_MT, mean_rel_happy)
```

```{r, echo=FALSE, error=FALSE}
uk %>%
  filter(!is.na(benefit_MT)) %>%
  filter(!is.na(rel_happy)) %>%
  ggplot(aes(
    wave,
    rel_happy,
    color = factor(benefit_MT,
                   labels = c("No means-tested benefit receipt", "Means-tested benefit receipt")),
    group = benefit_MT
  )) +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(
    y = "Relationship satisfaction",
    x = "Wave",
    color = "Means-tested benefit"
  ) 
```

### Moderation effects

```{r, echo=FALSE, error=FALSE}
uk %>%
  filter(!is.na(lfstat)) %>%
  filter(!is.na(benefit_MT)) %>%
  filter(!is.na(rel_happy)) %>%
  ggplot(aes(
    wave,
    rel_happy,
    color = factor(benefit_MT, labels = c("No means-tested benefit receipt", "Means-tested benefit receipt")),
    group = benefit_MT
  )) +
  geom_smooth(method = "lm") +
  facet_wrap(~ lfstat) +
  theme_bw() +
  labs(
    x = "Wave",
    y = "Relationship satisfaction",
    color = "Means-tested benefit"
  )


uk %>%
  filter(
    !is.na(job),
    !is.na(benefit_MT),
    !is.na(rel_happy)
  ) %>%
  ggplot(
    aes(
      x = wave,
      y = rel_happy,
      color = interaction(
        factor(
          benefit_MT,
          levels = c(0, 1),
          labels = c("No means-tested benefit receipt", "Means-tested benefit receipt")
        ),
        job,
        sep = " - "
      ),
      group = interaction(benefit_MT, job)
    )
  ) +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(
    x = "Wave",
    y = "Relationship satisfaction",
    color = "Benefit status and Paid Work (Last Week)"
  )
```


#### Dyadic Adjustment Scale: Relationship satisfaction

```{r, echo=FALSE}
uk %>%
  filter(!is.na(lfstat)) %>%
  filter(!is.na(benefit_MT)) %>%
  filter(!is.na(DAS_rel_sat)) %>%
  ggplot(aes(
    wave,
    DAS_rel_sat,
    color = factor(
      benefit_MT,
      labels = c(
        "No means-tested benefit receipt",
        "Means-tested benefit receipt"
      )
    ),
    group = benefit_MT
  )) +
  geom_smooth(method = "lm") +
  facet_wrap(~ lfstat) +
  theme_bw() +
  labs(
    x = "Wave",
    y = "Dyadic Adjustment Scale: Relationship satisfaction",
    color = "Means-tested benefit"
  )


uk %>%
  filter(
    !is.na(job),
    !is.na(benefit_MT),
    !is.na(DAS_rel_sat)
  ) %>%
  ggplot(
    aes(
      x = wave,
      y = DAS_rel_sat,
      color = interaction(
        factor(
          benefit_MT,
          levels = c(0, 1),
          labels = c(
            "No means-tested benefit receipt",
            "Means-tested benefit receipt"
          )
        ),
        job,
        sep = " - "
      ),
      group = interaction(benefit_MT, job)
    )
  ) +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(
    x = "Wave",
    y = "Dyadic Adjustment Scale: Relationship satisfaction",
    color = "Benefit status and Paid Work (Last Week)"
  )
```



























<!-- ## Statistics: Main variables -->

<!-- ```{r, echo=FALSE} -->
<!-- uk %>% -->
<!--   group_by(wave) %>% -->
<!--   summarise(Mean_HH_Net_Log_Income     = mean(log_hhnetinc_oecd, na.rm = T), -->
<!--             Mean_Benefit               = mean(benefit_MT, na.rm = T), -->
<!--             Mean_Rel_Happy             = mean(rel_happy, na.rm = T), -->
<!--             Mean_NChild                = mean(nchild, na.rm = T), -->
<!--             Mean_MCS                   = mean(mcs, na.rm = T), -->
<!--             Mean_PCS                   = mean(pcs, na.rm = T) -->
<!--   ) %>% -->
<!--   print(n = 16) -->


<!-- ``` -->

<!-- ### By LFS -->

<!-- ```{r, echo=FALSE, error=FALSE} -->
<!-- ### By LFS ---- -->
<!-- uk %>% -->
<!--   filter(!is.na(lfstat)) %>% -->
<!--   group_by(wave, lfstat) %>% -->
<!--   summarise(Mean_HH_Net_Log_Income = mean(log_hhnetinc_oecd, na.rm = T), -->
<!--             Mean_Benefit    = mean(benefit_MT, na.rm = T) -->
<!--   ) %>% -->
<!--   print(n = 90) -->
<!-- ``` -->

<!-- ### By paid work -->

<!-- ```{r, echo=FALSE, error=FALSE} -->
<!-- uk %>% -->
<!--   filter(!is.na(job)) %>% -->
<!--   group_by(wave, job) %>% -->
<!--   summarise(Mean_HH_Net_Log_Income = mean(log_hhnetinc_oecd, na.rm = T), -->
<!--             Mean_Benefit    = mean(benefit_MT, na.rm = T) -->
<!--   ) %>% -->
<!--   print(n = 30) -->
<!-- ``` -->

<!-- ### Means-tested benefit receipt only -->

<!-- ```{r, echo=FALSE, error=FALSE} -->
<!-- uk %>% -->
<!--   filter( -->
<!--     !is.na(lfstat), -->
<!--     benefit_MT == 1 -->
<!--   ) %>% -->
<!--   group_by(wave, lfstat) %>% -->
<!--   summarise( -->
<!--     Mean_HH_Net_Log_Income = mean(log_hhnetinc_oecd, na.rm = TRUE), -->
<!--     N_Benefit = n(), -->
<!--     .groups = "drop" -->
<!--   ) %>% -->
<!--   print (n = 90) -->
<!-- ``` -->

<!-- #### Benefit recipients and paid work -->

<!-- ```{r, echo=FALSE} -->
<!-- uk %>% -->
<!--   filter( -->
<!--     !is.na(job), -->
<!--     benefit_MT == 1 -->
<!--   ) %>% -->
<!--   group_by(wave, job) %>% -->
<!--   summarise( -->
<!--     Mean_HH_Net_Log_Income = mean(log_hhnetinc_oecd, na.rm = TRUE), -->
<!--     N_Benefit = n(), -->
<!--     .groups = "drop" -->
<!--   ) %>% -->
<!--   print (n = 90) -->
<!-- ``` -->













<!-- ## Correlation Matrices -->

<!-- ### Log Household Net Income OECD equivalence scale -->

<!-- ```{r, echo=FALSE} -->
<!-- income_cor <- uk %>% -->
<!--   select(pidp, wave, log_hhnetinc_oecd) %>% -->
<!--   arrange(pidp, wave) %>% -->
<!--   tidyr::pivot_wider( -->
<!--     names_from = wave, -->
<!--     values_from = log_hhnetinc_oecd, -->
<!--     names_prefix = "log_hhnetinc_oecd_" -->
<!--   ) %>% -->
<!--   select( -->
<!--     all_of( -->
<!--       paste0("log_hhnetinc_oecd_", 1:15) -->
<!--     ) -->
<!--   ) %>% -->
<!--   cor(use = "pairwise.complete.obs") %>% -->
<!--   round(2) -->

<!-- income_cor -->
<!-- ``` -->

<!-- ### Means-tested benefit receipt -->

<!-- ```{r, echo=FALSE} -->
<!-- benefitMT_cor <- uk %>% -->
<!--   select(pidp, wave, benefit_MT) %>% -->
<!--   tidyr::pivot_wider( -->
<!--     names_from = wave, -->
<!--     values_from = benefit_MT, -->
<!--     names_prefix = "benefit_MT_" -->
<!--   ) %>% -->
<!--   select( -->
<!--     any_of( -->
<!--       paste0("benefit_MT_", 1:15) -->
<!--     ) -->
<!--   ) %>% -->
<!--   cor(use = "pairwise.complete.obs") %>% -->
<!--   round(2) -->

<!-- benefitMT_cor -->
<!-- ``` -->

<!-- ### Relationship satisfaction -->

<!-- ```{r, echo=FALSE} -->
<!-- relhappy_cor <- uk %>% -->
<!--   select(pidp, wave, rel_happy) %>% -->
<!--   tidyr::pivot_wider( -->
<!--     names_from = wave, -->
<!--     values_from = rel_happy, -->
<!--     names_prefix = "rel_happy_" -->
<!--   ) %>% -->
<!--   select( -->
<!--     any_of( -->
<!--       paste0("rel_happy_", c(1, 3, 5, 7, 9, 11, 13, 15)) -->
<!--     ) -->
<!--   ) %>% -->
<!--   cor(use = "pairwise.complete.obs") %>% -->
<!--   round(2) -->

<!-- relhappy_cor -->
<!-- ``` -->










<!-- ## Income types -->

<!-- ### Houeshold Net Income -->

<!-- ```{r, echo=FALSE} -->
<!-- uk %>% -->
<!--   group_by(wave) %>% -->
<!--   summarise(N    = sum(!is.na(hhnetinc)), -->
<!--             Mean = mean(hhnetinc, na.rm = T), -->
<!--             Var  = var(hhnetinc, na.rm = T), -->
<!--             SD   = sd(hhnetinc, na.rm = T), -->
<!--             Min  = min(hhnetinc, na.rm = T), -->
<!--             Max  = max(hhnetinc, na.rm = T), -->
<!--             NAs  = sum(is.na(hhnetinc)) -->
<!--             ) -->
<!-- ``` -->

<!-- ### Correct high incomes? -->

<!-- ```{r, echo=FALSE} -->
<!-- uk %>% -->
<!--   filter(hhnetinc > 100000) %>% -->
<!--   select( -->
<!--     hidp, -->
<!--     wave, -->
<!--     hhnetinc, -->
<!--     hhgrsinc -->
<!--   ) %>% -->
<!--   distinct() %>% -->
<!--   arrange(desc(hhnetinc)) -->
<!-- ``` -->

<!-- ### Household Gross Income -->

<!-- ```{r, echo=FALSE} -->
<!-- uk %>% -->
<!--   group_by(wave) %>% -->
<!--   summarise(N    = sum(!is.na(hhgrsinc)), -->
<!--             Mean = mean(hhgrsinc, na.rm = T), -->
<!--             Var  = var(hhgrsinc, na.rm = T), -->
<!--             SD   = sd(hhgrsinc, na.rm = T), -->
<!--             Min  = min(hhgrsinc, na.rm = T), -->
<!--             Max  = max(hhgrsinc, na.rm = T), -->
<!--             NAs  = sum(is.na(hhgrsinc)) -->
<!--             ) -->
<!-- ``` -->

<!-- ### Household Net Income OECD equivalence scale -->

<!-- ```{r, echo=FALSE} -->
<!-- uk %>% -->
<!--   group_by(wave) %>% -->
<!--   summarise(N    = sum(!is.na(hhnetinc_oecd)), -->
<!--             Mean = mean(hhnetinc_oecd, na.rm = T), -->
<!--             Var  = var(hhnetinc_oecd, na.rm = T), -->
<!--             SD   = sd(hhnetinc_oecd, na.rm = T), -->
<!--             Min  = min(hhnetinc_oecd, na.rm = T), -->
<!--             Max  = max(hhnetinc_oecd, na.rm = T), -->
<!--             NAs  = sum(is.na(hhnetinc_oecd)) -->
<!--             ) -->
<!-- ``` -->

<!-- ### Household Gross Income OECD equivalence scale -->

<!-- ```{r, echo=FALSE} -->
<!-- uk %>% -->
<!--   group_by(wave) %>% -->
<!--   summarise(N    = sum(!is.na(hhgrsinc_oecd)), -->
<!--             Mean = mean(hhgrsinc_oecd, na.rm = T), -->
<!--             Var  = var(hhgrsinc_oecd, na.rm = T), -->
<!--             SD   = sd(hhgrsinc_oecd, na.rm = T), -->
<!--             Min  = min(hhgrsinc_oecd, na.rm = T), -->
<!--             Max  = max(hhgrsinc_oecd, na.rm = T), -->
<!--             NAs  = sum(is.na(hhgrsinc_oecd)) -->
<!--   ) -->
<!-- ``` -->

<!-- ### Log Household Net Income OECD equivalence scale -->

<!-- ```{r, echo=FALSE} -->
<!-- uk %>% -->
<!--   group_by(wave) %>% -->
<!--   summarise(N    = sum(!is.na(log_hhnetinc_oecd)), -->
<!--             Mean = mean(log_hhnetinc_oecd, na.rm = T), -->
<!--             Var  = var(log_hhnetinc_oecd, na.rm = T), -->
<!--             SD   = sd(log_hhnetinc_oecd, na.rm = T), -->
<!--             Min  = min(log_hhnetinc_oecd, na.rm = T), -->
<!--             Max  = max(log_hhnetinc_oecd, na.rm = T), -->
<!--             NAs  = sum(is.na(log_hhnetinc_oecd)) -->
<!--             ) -->
<!-- ``` -->

<!-- ### Log Household Gross Income OECD equivalence scale -->

<!-- ```{r, echo=FALSE} -->
<!-- uk %>% -->
<!--   group_by(wave) %>% -->
<!--   summarise(N    = sum(!is.na(log_hhgrsinc_oecd)), -->
<!--             Mean = mean(log_hhgrsinc_oecd, na.rm = T), -->
<!--             Var  = var(log_hhgrsinc_oecd, na.rm = T), -->
<!--             SD   = sd(log_hhgrsinc_oecd, na.rm = T), -->
<!--             Min  = min(log_hhgrsinc_oecd, na.rm = T), -->
<!--             Max  = max(log_hhgrsinc_oecd, na.rm = T), -->
<!--             NAs  = sum(is.na(log_hhgrsinc_oecd)) -->
<!--             ) -->
<!-- ``` -->



<!-- ##### Summary -->

<!-- ```{r, echo=FALSE} -->
<!-- income_summary <- uk %>% -->
<!--   summarise( -->
<!--     across( -->
<!--       c( -->
<!--         hhnetinc, -->
<!--         hhgrsinc, -->
<!--         hhnetinc_oecd, -->
<!--         hhgrsinc_oecd, -->
<!--         log_hhnetinc_oecd, -->
<!--         log_hhgrsinc_oecd -->
<!--       ), -->
<!--       list( -->
<!--         N = ~sum(!is.na(.)), -->
<!--         Mean = ~mean(., na.rm = TRUE), -->
<!--         SD = ~sd(., na.rm = TRUE), -->
<!--         Min = ~min(., na.rm = TRUE), -->
<!--         P25 = ~quantile(., 0.25, na.rm = TRUE), -->
<!--         Median = ~median(., na.rm = TRUE), -->
<!--         P75 = ~quantile(., 0.75, na.rm = TRUE), -->
<!--         P95 = ~quantile(., 0.95, na.rm = TRUE), -->
<!--         P99 = ~quantile(., 0.99, na.rm = TRUE), -->
<!--         Max = ~max(., na.rm = TRUE) -->
<!--       ), -->
<!--       .names = "{.col}_{.fn}" -->
<!--     ) -->
<!--   ) %>% -->
<!--   pivot_longer( -->
<!--     everything(), -->
<!--     names_to = c("Variable", ".value"), -->
<!--     names_sep = "_(?=[^_]+$)" -->
<!--   ) -->
<!-- #income_summary -->

<!-- income_summary %>% -->
<!--   mutate( -->
<!--     Variable = recode( -->
<!--       Variable, -->
<!--       hhnetinc = "Household net income", -->
<!--       hhgrsinc = "Household gross income", -->
<!--       hhnetinc_oecd = "Equivalised household net income", -->
<!--       hhgrsinc_oecd = "Equivalised household gross income", -->
<!--       log_hhnetinc_oecd = "Log equivalised household net income", -->
<!--       log_hhgrsinc_oecd = "Log equivalised household gross income" -->
<!--     ) -->
<!--   ) %>% -->
<!--   gt() %>% -->
<!--   fmt_integer( -->
<!--     columns = N, -->
<!--     use_seps = TRUE -->
<!--   ) %>% -->
<!--   fmt_number( -->
<!--     columns = c(Mean, SD, Min, P25, Median, P75, P95, P99, Max), -->
<!--     decimals = 2, -->
<!--     use_seps = TRUE -->
<!--   ) %>% -->
<!--   cols_label( -->
<!--     Variable = "Income measure", -->
<!--     N = "N", -->
<!--     Mean = "Mean", -->
<!--     SD = "SD", -->
<!--     Min = "Min", -->
<!--     P25 = "P25", -->
<!--     Median = "Median", -->
<!--     P75 = "P75", -->
<!--     P95 = "P95", -->
<!--     P99 = "P99", -->
<!--     Max = "Max" -->
<!--   ) %>% -->
<!--   tab_header( -->
<!--     title = "Household Income Statistics" -->
<!--   ) -->
<!-- ``` -->

<!-- ##### Summary by wave -->

<!-- ```{r, echo=FALSE} -->
<!-- income_summary_wave <- uk %>% -->
<!--   group_by(wave) %>% -->
<!--   summarise( -->
<!--     across( -->
<!--       c( -->
<!--         hhnetinc, -->
<!--         hhgrsinc, -->
<!--         hhnetinc_oecd, -->
<!--         hhgrsinc_oecd, -->
<!--         log_hhnetinc_oecd, -->
<!--         log_hhgrsinc_oecd -->
<!--       ), -->
<!--       list( -->
<!--         N = ~sum(!is.na(.x)), -->
<!--         Mean = ~mean(.x, na.rm = TRUE), -->
<!--         SD = ~sd(.x, na.rm = TRUE), -->
<!--         Min = ~min(.x, na.rm = TRUE), -->
<!--         P25 = ~quantile(.x, 0.25, na.rm = TRUE), -->
<!--         Median = ~median(.x, na.rm = TRUE), -->
<!--         P75 = ~quantile(.x, 0.75, na.rm = TRUE), -->
<!--         P95 = ~quantile(.x, 0.95, na.rm = TRUE), -->
<!--         P99 = ~quantile(.x, 0.99, na.rm = TRUE), -->
<!--         Max = ~max(.x, na.rm = TRUE) -->
<!--       ), -->
<!--       .names = "{.col}_{.fn}" -->
<!--     ), -->
<!--     .groups = "drop" -->
<!--   ) %>% -->
<!--   pivot_longer( -->
<!--     cols = -wave, -->
<!--     names_to = c("variable", ".value"), -->
<!--     names_pattern = "^(.*)_(N|Mean|SD|Min|P25|Median|P75|P95|P99|Max)$" -->
<!--   ) %>% -->
<!--   arrange(wave, variable) -->

<!-- # income_summary_wave %>% -->
<!-- #   print (n = 91) -->

<!-- income_summary_wave %>% -->
<!--   mutate( -->
<!--     variable = recode( -->
<!--       variable, -->
<!--       hhnetinc = "Household net income", -->
<!--       hhgrsinc = "Household gross income", -->
<!--       hhnetinc_oecd = "Equivalised household net income", -->
<!--       hhgrsinc_oecd = "Equivalised household gross income", -->
<!--       log_hhnetinc_oecd = "Log equivalised household net income", -->
<!--       log_hhgrsinc_oecd = "Log equivalised household gross income" -->
<!--     ) -->
<!--   ) %>% -->
<!--   gt() %>% -->
<!--   fmt_integer( -->
<!--     columns = N, -->
<!--     use_seps = TRUE -->
<!--   ) %>% -->
<!--   fmt_number( -->
<!--     columns = c(Mean, SD, Min, P25, Median, P75, P95, P99, Max), -->
<!--     decimals = 2, -->
<!--     use_seps = TRUE -->
<!--   ) %>% -->
<!--   cols_label( -->
<!--     wave = "Wave", -->
<!--     variable = "Income measure", -->
<!--     N = "N", -->
<!--     Mean = "Mean", -->
<!--     SD = "SD", -->
<!--     Min = "Min", -->
<!--     P25 = "P25", -->
<!--     Median = "Median", -->
<!--     P75 = "P75", -->
<!--     P95 = "P95", -->
<!--     P99 = "P99", -->
<!--     Max = "Max" -->
<!--   ) %>% -->
<!--   tab_header( -->
<!--     title = "Household Income Statistics by Wave" ) %>% -->
<!--    tab_options( -->
<!--     table.font.size = px(11), -->
<!--     data_row.padding = px(2), -->
<!--     heading.title.font.size = px(14), -->
<!--     column_labels.font.size = px(11) -->
<!--   ) -->
<!-- ``` -->













<!-- ### Sankey/river/alluvial plot -->

<!-- ```{r, echo=FALSE, error=FALSE, include=FALSE} -->
<!-- trans_data <- uk %>% -->
<!--   select(pidp, wave, benefit_MT) %>% -->
<!--   tidyr::pivot_wider( -->
<!--     names_from = wave, -->
<!--     values_from = benefit_MT, -->
<!--     names_prefix = "benefit_MT_" -->
<!--   ) %>% -->
<!--   select( -->
<!--     pidp, -->
<!--     any_of(paste0("benefit_MT_", 1:15)) -->
<!--   ) %>% -->
<!--   count( -->
<!--     across(starts_with("benefit_MT_")) -->
<!--   ) %>% -->
<!--   mutate( -->
<!--     id = row_number() -->
<!--   ) -->

<!-- trans_data_long <- trans_data %>% -->
<!--   gather( -->
<!--     key = key, -->
<!--     value = value, -->
<!--     -n, -->
<!--     -id -->
<!--   ) -->

<!-- trans_data_long <- trans_data_long %>% -->
<!--   mutate( -->
<!--     wave = str_remove(key, "benefit_MT_") %>% -->
<!--       as.numeric(), -->

<!--     value = if_else( -->
<!--       is.na(value), -->
<!--       "Missing", -->
<!--       as.character(value) -->
<!--     ), -->

<!--     value = factor(value), -->

<!--     value = fct_relevel( -->
<!--       value, -->
<!--       "0", "1", "Missing" -->
<!--     ) -->
<!--   ) -->

<!-- plot <- ggplot( -->
<!--   trans_data_long, -->
<!--   aes( -->
<!--     x = wave, -->
<!--     y = n, -->
<!--     stratum = value, -->
<!--     fill = value, -->
<!--     alluvium = id -->
<!--   ) -->
<!-- ) + -->
<!--   geom_stratum(alpha = .5) + -->
<!--   geom_flow() -->
<!-- ``` -->

<!-- ```{r, echo=FALSE, error=FALSE} -->
<!-- plot + -->
<!--   theme_tufte(base_size = 18) + -->
<!--   labs( -->
<!--     x = "Wave", -->
<!--     y = "Frequency", -->
<!--     fill = "means-tested benefit receipt" -->
<!--   ) + -->
<!--   scale_fill_viridis_d(direction = -1) -->
<!-- ``` -->
